Jane, John … Leslie? A Historical Method for Algorithmic Gender Prediction

### 2015-08-21 article.Rmd
Fix footnote

As it currently stands, the gender package meets a methodological need. As of August 2015, the gender package has been downloaded more than 21,000 times and incorporated as a pedagogical tool within digital humanities courses [Jockers 2014b].^[The download counts come from the [download logs from RStudio's CRAN mirror](http://cran-logs.rstudio.com/) as tabulated by Gabor Csardi's [cranlogs package](http://cranlogs.r-pkg.org/). [Csardi 2015] RStudio is the only CRAN mirror to release its logs.) Researchers have used it to uncover instances gender inequity in a range of different areas, from biases in students' teaching evaluations to disparities between attendees and authors at the annual Digital Humanities Conference [Schmidt 2015, Weingart 2015]. These implementations help illustrate the value of the method. Its major intervention is to algorithmically inferring gender in textual sources in a way that takes into account change over time. Rather than trying to shoehorn existing computer science methods to fit humanities research, the gender package reverses this traditional disciplinary relationship. It uses a humanistic starting point---the historical relationship between gender and naming practices---to contribute a new method to computer science.

>>>>

As it currently stands, the gender package meets a methodological need. As of August 2015, the gender package has been downloaded more than 21,000 times and incorporated as a pedagogical tool within digital humanities courses [Jockers 2014b].^[The download counts come from the [download logs from RStudio's CRAN mirror](http://cran-logs.rstudio.com/) as tabulated by Gabor Csardi's [cranlogs package](http://cranlogs.r-pkg.org/). [Csardi 2015] RStudio is the only CRAN mirror to release its logs.)] Researchers have used it to uncover instances gender inequity in a range of different areas, from biases in students' teaching evaluations to disparities between attendees and authors at the annual Digital Humanities Conference [Schmidt 2015, Weingart 2015]. These implementations help illustrate the value of the method. Its major intervention is to algorithmically inferring gender in textual sources in a way that takes into account change over time. Rather than trying to shoehorn existing computer science methods to fit humanities research, the gender package reverses this traditional disciplinary relationship. It uses a humanistic starting point---the historical relationship between gender and naming practices---to contribute a new method to computer science.
